---
title: "CUNY 605"
subtitle: "Week 10 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

m2l <- function(matr) {
  ### Convert R matrix to Latex
  printmrow <- function(x) {
    ret <- paste(paste(x,collapse = " & "),"\\\\")
    sprintf(ret)
  }
  out <- apply(matr,1,printmrow)
  out2 <- paste("\\begin{bmatrix}",paste(out,collapse=' '),"\\end{bmatrix}")
  return(out2)
}
```

<br>

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. 
 
Find the probability that he wins 8 dollars before losing all of his money if 
 
 
(a) he bets 1 dollar each time (timid strategy)
(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy)
(c) Which strategy gives Smith the better chance of getting out of jail?

<br>

###Part (a)

Solve using markov chain matrix absorption:

```{r mcm}
### Define states
state_labels <- c("broke", 1:7, "free")

### Set up transition matrix
up_prob <- 0.4
down_prob <- 0.6
tr_mat <- matrix(0, dimnames = list(state_labels, state_labels), nrow=length(state_labels), ncol=length(state_labels))
tr_mat[1,1] <- 1
tr_mat[length(state_labels),length(state_labels)] <- 1
diag(tr_mat[-length(state_labels), -1]) <- c(0, rep(up_prob, length(state_labels)-2))
diag(tr_mat[-1, -length(state_labels)]) <- c(rep(down_prob, length(state_labels)-2), 0)

### Decompose transition matrix into submatrices
t <- state_labels[-c(1, length(state_labels))]
r <- c(state_labels[1], state_labels[length(state_labels)])
I <- tr_mat[r,r]
Q <- tr_mat[t,t]
R <- tr_mat[t,r]
O <- tr_mat[r,t]

### Find expected number of times in each transient state (columns) if starting in a transient state (rows)
N <- solve(diag(nrow(Q)) - Q)

### Find expected number of times until absorption (column) if starting in a transient state (rows)
t_abs <- N %*% cbind(rep(1, nrow(N)))

### Find probability of absorption (columns) if starting in a transient state (rows)
B <- N %*% R

B["1", ]
```

Compare to empirical answer of multiplying initial state matrix by transition matrix n->inf times

```{r emp}
### Set up initial state matrix
in_mat <- matrix(0, nrow=1, ncol=length(state_labels), dimnames = list("initial", state_labels))
in_mat[, 2] <- 1

### Write function to take a matrix to any power
mat_power <- function(mat, n){
  new_mat <- mat
  flag <- n-1
  while (flag>0) {
    new_mat <- new_mat %*% mat
    flag <- flag - 1
  }
  return(new_mat)
}

in_mat %*% mat_power(tr_mat, 100000)

```

Compare to Gambler's Ruin formula

```{r gr}
i <- 1
N <- 8
p <- 0.4
q <- 1-p

free <- (1-(q/p)^i) / (1-(q/p)^N)
broke <- 1 - free

broke; free
```

Compare to Monte Carlo simulation

```{r monte}
calc_final_state <- function(money=1, up_prob=0.4, down_prob=0.6){
  while (money > 0 & money < 8){
    money <- money + sample(c(-1,1), 1, prob=c(down_prob, up_prob), replace=T)
  }
  return(money)
}

results <- replicate(100000, calc_final_state())
table(results)/length(results)

```

There is more statistical error in the simulation, but all answers are similar.

<br>

###Part (b)

This is just the equivalent of winning three in a row, i.e. 0.4^3 = 0.064

<br>

###Part (c)

The strategy in (b) gives Smith a better chance of getting out of jail.
