---
title: "CUNY 605"
subtitle: "Week 15 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

###1

Regression equation

```{r ex_1}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)
lm_fit <- lm(y ~ x)
summary(lm_fit)
```

$$ y = `r round(summary(lm_fit)$coefficients[1,1],2)` + `r round(summary(lm_fit)$coefficients[2,1],2)` x $$

###2

Critical points (minima, maxima, saddle points)

$$ f(x, y) = 24x - 6xy^2 - 8y^3 $$

Find partial derivatives:

$$ f_x(x, y) = 24 - 6y^2 $$
$$ f_y(x, y) = -12xy - 24y^2 $$
Set both equal to zero and solve.

$$ y^2 = 4 ; y = \pm 2 $$
$$ x = -4 $$
Plug into original equation to get:  

__(-4, 2, -64) and (-4, -2, 64)__

###3

__Step One__  

Take the two formulas for units sold, multiply them each by dollars, and add them.

$$ revenue(x, y) = x(81 - 21x + 17y) + y(40 + 11x - 23y) $$
$$ = 81x - 21x^2 + 17xy + 40y + 11xy - 23y^2 $$
$$ = -21x^2 - 23y^2 + 28xy + 81x + 40y $$

__Step Two__  

$$ revenue(2.3, 4.1) = 116.62 $$

###4

Minimize costs by finding the partial derivatives of the cost function, then setting them equal to each other, and substituting back into the constraint formula.

$$ C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700 $$
$$ C_x(x,y) = \frac{1}{3} x + 7 $$
$$ C_y(x,y) = \frac{1}{3} y + 25 $$
$$ \frac{1}{3} x + 7 = \frac{1}{3} y + 25 $$
$$ x = y + 54 $$

Constraint formula:

$$ x + y = 96 $$
$$ y + 54 + y = 96 $$
$$ y = 21 $$
$$ x = 75 $$

###5

Double integral

$$ \int_2^4 \int_2^4 (e^{8x + 3y}) dy dx $$
$$ = \int_2^4 ([\frac{1}{3}e^{8x + 3y}] \rvert_2^4) dx $$
$$ = \int_2^4 (\frac{1}{3}e^{8x + 12} - \frac{1}{3}e^{8x + 6}) dx $$
$$ = \int_2^4 (\frac{1}{3}e^{8x}(e^{12} - e^{6})) dx $$
$$ = (\frac{1}{24}e^{8x}(e^{12} - e^{6})) \rvert_2^4 $$
$$ = (\frac{1}{24}e^{32}(e^{12} - e^{6}) - \frac{1}{24}e^{16}(e^{12} - e^{6})) $$
= __534,155,900,000,000,000__
