---
title: "CUNY 605"
subtitle: "Week 4 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

m2l <- function(matr) {
  ### Convert R matrix to Latex
  printmrow <- function(x) {
    ret <- paste(paste(x,collapse = " & "),"\\\\")
    sprintf(ret)
  }
  out <- apply(matr,1,printmrow)
  out2 <- paste("\\begin{bmatrix}",paste(out,collapse=' '),"\\end{bmatrix}")
  return(out2)
}
```

###Problem Set 1

Singular Value Decomposition: The goal of "diagonalizing" a matrix is to find an orthonormal basis that gets transformed to another orthonormal basis (times some scalar) after applying the linear transformation (i.e., multiplying by the matrix.)  

In other words, if we have a matrix A, we are looking for A V = U D where V and U are matrices of orthonormal basis vectors and D is a diagonal matrix of "scalars."  

Rather than trying to find V and U simultaneously, we can use algebra to arrive at the following formula:   
A^T^ A = V D^2^ V^T^  
where V is the eigenvectors and D^2^ is the eigenvalues of the matrix A^T^ A (and we can then just take the square root of D^2^ to get the D scalar values we were originally looking for.)  
Likewise, A A^T^ = U D^2^ U^T^

```{r ps_1}

# To test, we start with a matrix A
A <- rbind(c( 1, 2, 3),
           c(-1, 0, 4))

# Then we create AA^t and A^tA
X <- A %*% t(A)

Y <- t(A) %*% A

# Using R's eigen() function we can get the eigenvectors and eigenvalues of X and Y
X_eig <- eigen(X)

Y_eig <- eigen(Y)

# And we compare the output to the svd() function applied to our original matrix A
A_svd <- svd(A)

# The square root of the eigenvalues of X (and Y too) should be the same as the D in our svd of A:
sqrt(X_eig$values)
sqrt(Y_eig$values)
A_svd$d

# The eigenvectors of X should be the same (or same span) as U
X_eig$vectors
A_svd$u

# The eigenvectors of Y should be the same (or same span) as V
Y_eig$vectors
A_svd$v

```


###Problem Set 2

Create a function that computes the inverse of a matrix using co-factors.

```{r ps_2}

myinverse <- function(A){
  C <- array(0, dim=dim(A))
  for (i in 1:nrow(A)){
    for (j in 1:ncol(A)){
      C[i,j] <- (-1)^(i+j)*det(A[-i,-j])
    }
  }
  t(C)/det(A)
}

# Test the function
A <- rbind(c(1,2,3),
           c(7,0,4),
           c(4,9,0))

B <- myinverse(A)

# Compare to solve() function
B
solve(A)

# AB=I
round(A %*% B, 2)

```

