---
title: "CUNY 605"
subtitle: "Week 11 Discussion"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

<br>

__For this example we will get Major League Baseball data from espn.com and fit a simple linear regression.__

<br>

###Libraries

First we load the necessary libraries.

```{r library_load}
library(dplyr)
library(rvest)
library(broom)
library(ggplot2)
library(stringr)
library(knitr)
```

<br>

###Get the data

Then use rvest to scrape the 2018 MLB standings from espn.com and clean up the team abbreviations using regular expressions.

```{r get_data}
season <- 2018
url <- paste0('http://www.espn.com/mlb/standings/_/sort/winpercent/dir/desc/season/', season, '/group/overall')
url_html <- url %>% read_html()
html_tables <- url_html %>% html_table(fill=T)
teams <- html_tables[[2]]

# use regex to extract team acronym - get uppercase letters followed by lowercase, then delete last two characters
teams <- teams %>% rename(team=1) %>% mutate(team=str_extract_all(team, "[A-Z]{3,4}[a-z]") %>% substr(1, nchar(.)-2))

standings <- html_tables[[4]]

team_standings <- bind_cols(teams, standings)
team_standings %>% head() %>% kable()
```

<br>

###Build the model

Run differential should correlate to to winning. We examine a scatterplot and see that a simple linear model looks appropriate.

```{r build_model}
team_standings %>% 
  ggplot(aes(x=DIFF, y=W)) + 
  geom_point() + 
  geom_smooth(method="lm", se=T) + 
  labs(x="Run Differential", y="Number of Wins")
lm_fit <- lm(W ~ DIFF, data=team_standings)
summary(lm_fit)
```

The high adjusted R-squared indicates a good fit.

<br>

###Examine the residuals

We use the broom package to put the linear model into a data frame.

```{r residuals}
lm_df <- augment(lm_fit)
lm_df %>% 
  ggplot(aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  geom_hline(yintercept = 0)
```

The residual values appear randomly scattered, as they should in a well fit model.
