---
title: "CUNY 605"
subtitle: "Week 2 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
m2l <- function(matr) {
    
    
    
    printmrow <- function(x) {
        
        
        
        ret <- paste(paste(x,collapse = " & "),"\\\\")
        
        sprintf(ret)
        
    }
    
    
    
    out <- apply(matr,1,printmrow)
    
    out2 <- paste("\\begin{bmatrix}",paste(out,collapse=' '),"\\end{bmatrix}")
    
    return(out2)
    
}
```

###Problem Set 1

(1) A^T^A does not equal AA^T^

Assume the following two matrices:

$$\mathbf{A} = \left[\begin{array}
{rrr}
A_{11} & A_{12} \\ 
A_{21} & A_{22}
\end{array}\right]
$$

$$\mathbf{A^T} = \left[\begin{array}
{rrr}
A_{11} & A_{21} \\ 
A_{12} & A_{22}
\end{array}\right]
$$
If we multiply AA^T^ the value in the first cell (position 1,1) would be
$$ A_{11}A_{11}+A_{12}A_{12} $$
whereas if we multiply A^T^A the value in the first cell would be
$$A_{11}A_{11}+A_{21}A_{21}$$
So clearly the two are not equivalent.

(2) A^T^A = AA^T^ only when A is symmetric. Using the above proof as an example, A~12~ would be the same as A~21~ if the matrix were symmetric.


###Problem Set 2

```{r problem_2}
LU_decomp <- function(A){
  
  ### Check to make sure matrix is not singular, rectangular, or comprised of zeros on pivot
  if(sum(diag(A)==0)>0 | qr(A)$rank<max(nrow(A), ncol(A))) {
    message("Error: This matrix cannot be decomposed.")
  } else {
    U <- A
    # Get U by going column by column and algebraically putting zeros in all positions below the diagonal
    for (j in 1:(ncol(U)-1)){
      for (i in (j+1):nrow(U)){
        if (U[i,j]!=0){U[i,] <- U[i, ] - U[j, ]*(U[i,j]/U[j,j])}
      }
    }
    
    # Get L by restating L U = A  as  U_inv L_inv = A_inv  ,solving for L_inv, and then taking inverse to get L 
    U_inv <- solve(U)
    A_inv <- solve(A)
    L_inv <- solve(U_inv, A_inv)
    L <- solve(L_inv)
    list(L=round(L,2) , U=round(U,2))
  }
}

### Test the function
test_A <- rbind(c(1, 1, 3),
                c(2, -1, 5),
                c(-1, -2, 4))
LU <- LU_decomp(test_A)

```

$$A= `r m2l(test_A)` $$

$$L= `r m2l(LU[["L"]])` $$

$$U= `r m2l(LU[["U"]])` $$
```{r bad_matrices}
# try matrix with zero pivot
test_A <- rbind(c(0,1,1),
                c(1,2,3),
                c(7,7,9))
LU_decomp(test_A)

# try matrix that is singular
test_A <- rbind(c(1,1,1),
                c(1,2,3),
                c(7,7,7))
LU_decomp(test_A)