---
title: "CUNY 605"
subtitle: "Week 13 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

###1. Integration by substitution

$$ \int 4 e^{-7 x} d x $$
u = -7x  
du/dx = -7  
du = -7 dx  
$$ = \int \frac{-4}{7} d u e^u $$
$$ = \frac{-4}{7} \int e^u d u $$
$$ = \frac{-4}{7} e^u + C $$
$$ = \frac{-4}{7} e^{-7x} + C $$

###2. Integration of 1/x^n^

$$ \int -\frac{3150}{t^4} - 220 $$
$$ = \int -3150 t^{-4} - 220 $$
$$ = -3150 \frac{t^{-3}}{-3} - 220 t + C = 6530 $$
$$ = 1050 t^{-3} - 220 t + 5700 $$

###3. Area of rectangles

1 + 3 + 5 + 7 = __16__

###4. Area between functions

```{r ex_4}
curve(x^2 - 2*x - 2, -5, 5)
curve(x + 2, -5, 5, add=T, col="red")
```

Set the equations equal to each other to calculate where they intersect.

x^2 - 2*x - 2 = x + 2  
x^2 - 3*x - 4 = 0  
(x-4)*(x+1) = 0  
x = (4, -1)  

$$ \int_{-1}^4 (x + 2) - (x^2 - 2x -2) $$
$$ = \int  -x^2 + 3x + 4 $$

$$ = \frac{x^3}{3} + \frac{3 x^2}{2} + 4 x + C $$
Plugging in 4 and -1 and subtracting:

4^3/3 + 3\*4^2/2 + 4\*4 - ((-1)^3/3 + 3\*(-1)^2/2 + 4\*(-1)) = __`r 4^3/3 + 3*4^2/2 + 4*4 - ((-1)^3/3 + 3*(-1)^2/2 + 4*(-1))`__

###5. Business example

Unless there is information missing (no price or demand curve formula is given), selling all 110 irons in one order would minimize costs.

###6. Integration by Parts

$$ \int ln(9x) x^6 dx $$
$$ \int f(x) g'(x) dx = f(x) g(x) - \int f'(x) g(x) dx $$
f(x) = ln(9x)  
f'(x) = 1/x  
g'(x) = x^6^  
g(x) = x^7^/7  
$$ ln(9x) \frac{x^7}{7} - \int \frac{1}{x} \frac{x^7}{7} dx $$
$$ ln(9x) \frac{x^7}{7} - \frac{1}{7} \int x^6 dx $$
$$ ln(9x) \frac{x^7}{7} - \frac{1}{7} \frac{x^7}{7} + C $$
$$ \frac{x^7}{49}(7 ln(9x) - 1)  + C $$

###7. Definite Integral

$$ \int_1^{e^6} \frac{1}{6x} dx $$
$$ \frac{1}{6} \int x^{-1} dx $$
$$ \frac{ln x}{6} $$
Now plugging in e^6^ and 1 and subtracting:

log(exp(6))/6 - log(exp(1))/6 = __`r log(exp(6))/6 - log(exp(1))/6`__

It is not a probability distribution because the area is not 1.
